---
author: "Julia Ran, William Parker"
output:
  rmdformats::material:
    highlight: tango
  html_document:
    theme: cosmo
    toc: yes
---

# Loading in packages 

This chunk is used to load in the packages in R that help us to tidy, manipulate, and visualize the data. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

require("knitr")
opts_knit$set(root.dir = "C:/Users/julia/SRP/SRP/Codes/SAF SAS files")

```


```{r library}
library("ggplot2")  
library("tidyverse")
library("haven")
library("rmdformats")
library("ggthemes")
library("dplyr")
library("tidyselect")
library(nnet)
library(rmdformats)
library(tableone)
```


## Loading in Final_Sample file 

The file being uploaded is the same file that was the output of the data_preparation rmd. 
```{r read_in_final_sample}
setwd("C:/Users/julia/SRP/SRP/Codes/SAF SAS files")

final_sample <- read.csv("C:/Users/julia/SRP/SRP/Codes/SAF SAS files/final_sample.csv")

```

## Creating functions

We created some functions in order to make our code more concise. 
```{r two_rates_with_diff}
comma <- function(x){
  case_when(
    abs(x) > 10 ~ format(x, digits = 0, big.mark = ","),
    abs(x) > 1 ~ format(x, digits = 2, big.mark = ","),
    TRUE ~ format(x, digits = 2, big.mark = ",")
  )
    
}

comma_p <- function(x){
  case_when(
    x < 0.001 ~ "p<0.001",
    x < 0.01 ~ paste0("p=" ,format(x, digits = 1, big.mark = ",")),
    TRUE ~ paste0("p=" ,format(x, digits = 2, big.mark = ","))
  )
}

rates_by_policy_diff <- function(name, final_sample){
  df <- final_sample %>%
    group_by(policy) %>%
    filter(!is.na(policy)) %>%
    count(status) %>% 
    mutate(total = sum(n),
           pct_status = 100*n/total) %>%
    filter(status == name)
  
  pre <- df[1,]
  post <- df[2,]
  
  a <- prop.test(x = c(post$n, pre$n), n = c( post$total_treated, pre$total_treated))
  low_ci <- 100*a$conf.int[1]
  
  up_ci <- 100*a$conf.int[2]
  post <- df$percent_treated[[2]]
  pre <- df$percent_treated[[1]]
  diff <- (post - pre)
  #paste0(name, "%; 95% CI ", comma(low_ci), "% to ", comma(up_ci), "%).")
  paste0(name, " was used to list ", comma(pre), "% of candidates in the 3-tier system to ", comma(post), "% candidates in the 6-tier system (absolute difference ", comma(diff), "%; 95% CI ", comma(low_ci), "% to ", comma(up_ci), "%).")
  
}
paragraph <- ""

for (name in levels(final_sample$status)){
  paragraph <- paste0(paragraph, " ", rates_by_policy_diff(name, final_sample))
}
```

since there is no status 5 patients in the pre-policy group, I deselecte them to help the code run more smoothly.

# Table 1: Candidate demographics and clinical characteristics by policy period
```{r select_table_one_variables, warning = FALSE}
setwd("C:/Users/julia/SRP/SRP/Codes/SAF SAS files")
cand_thor <- haven::read_sas("cand_thor.sas7bdat", NULL)# loading in "cand_thor" 

for_table_1 <- final_sample %>% 
  filter(is.na(policy) == FALSE) %>%
  filter(status != "Status 5") %>% 
  dplyr::select(PX_ID,policy,treatment) %>%
  left_join(cand_thor) %>%
  filter(PX_ID %in% final_sample$PX_ID) %>%
   mutate(age = CAN_AGE_AT_LISTING,
         female = ifelse(CAN_GENDER == "F", "1", "0"),
         race = factor(CAN_RACE),
         race = fct_lump(race, n = 3),
         race = fct_recode(race,
                          "White" = "8",
                          "Black" = "16",
                          "Hispanic" = "2000", 
                          "Other" = "Other"),
         bmi = CAN_BMI,
        simple_diagnosis = case_when(
          CAN_DGN>999 & CAN_DGN<1007 ~ "Dilated cardiomyopathy, non-ischemic",
          CAN_DGN == 1007 | CAN_DGN ==1200 ~ "Ischemic cardiomyopathy",
          CAN_DGN>1048 & CAN_DGN< 1100 ~ "Restrictive cardiomyopathy",
          TRUE ~ "Other"
        ),
        "Diagnosis" = factor(simple_diagnosis, 
                           levels = c("Dilated cardiomyopathy, non-ischemic", 
                                      "Ischemic cardiomyopathy", 
                                      "Restrictive cardiomyopathy", 
                                      "Other")),
        diabetes = case_when(
          CAN_DIAB_TY>1 & CAN_DIAB_TY<6 ~ "History of DM",
          CAN_DIAB_TY ==1 ~ "Non-diabetic",
          TRUE ~ "Unknown"
        ),
        diabetes = factor(diabetes),
        female_gfr = if_else(CAN_GENDER == "F", 0.742, 1),
        black_gfr = if_else(race == "Black", 1.21, 1),
        eGFR = 175*((CAN_MOST_RECENT_CREAT)^(-1.154))*(CAN_AGE_AT_LISTING^(-0.203))*female_gfr*black_gfr,
        functional_status = case_when(
          CAN_FUNCTN_STAT == 1 | (CAN_FUNCTN_STAT>2069) ~
            "Limited Impairment, 100-70%",
          CAN_FUNCTN_STAT == 2 | (CAN_FUNCTN_STAT>2049 & CAN_FUNCTN_STAT<2061) ~ 
            "Moderate Impairment, 50-60%",
          CAN_FUNCTN_STAT == 3 | (CAN_FUNCTN_STAT>2000 & CAN_FUNCTN_STAT<2041) ~ as.character(expression("Severe Impairment,  \u2264 40%")),
          TRUE ~ "Unknown"
        ),
        functional_status = ifelse(is.na(functional_status), "Unknown", functional_status),
        functional_status = factor(functional_status),
        pcwp = CAN_PCW_MEAN,
        body_surface_area = 0.007184*(CAN_HGT_CM)^(0.725)*CAN_WGT_KG^(0.425),
        cardiac_index = as.numeric(CAN_CARDIAC_OUTPUT/body_surface_area),
        cardiac_index = ifelse(cardiac_index>10, NA, cardiac_index),
        blood_type = factor(
           case_when(
             CAN_ABO %in% c("A", "A1", "A2") ~ "A",
             CAN_ABO %in% c("A1B", "A2B") ~ "AB",
             TRUE ~ CAN_ABO)
           ),
        payor = case_when(
          CAN_PRIMARY_PAY %in% c(3,4,13) ~ "Medicare",
          CAN_PRIMARY_PAY ==2 ~ "Medicaid",
          CAN_PRIMARY_PAY == 1 ~ "Private",
          TRUE ~ "Other"
        ),
        History_of_Smoking = case_when(
           CAN_HIST_CIGARETTE == "Y" ~ "Smoking history",
           CAN_HIST_CIGARETTE == "N" ~ "No prior smoking history"),
        Working = case_when(
           CAN_WORK_INCOME == "N" ~ "Not working",
           CAN_WORK_INCOME == "Y" ~ "Working"),
        Education_Status = case_when(
          CAN_EDUCATION %in% c(4,5,6) ~ "College",
          CAN_EDUCATION == 3 ~ "High School",
          TRUE ~ "Less than high school or other"
        ),
  ) %>%
  dplyr::select(PX_ID, policy, #patient ID and listing status
         age, female, race, #basic demographics
         History_of_Smoking, Working, Education_Status,
         bmi, blood_type, "Diagnosis", eGFR, diabetes, functional_status, #diagnosis and medical comorbidities, some factor variables some continuous
         cardiac_index, pcwp, #hemodynamics
         payor, treatment #payor
         #ecmo, iabp, lvad, other_mcs, high_dose_inotropes, low_dose_inotropes #treatments- hot coded
         ) %>%
  mutate_if(is.character, factor) %>% 
  mutate(bmi = substr(bmi, 0, 2)) %>% 
  mutate(bmi = as.numeric(bmi))

haven::write_dta(for_table_1, "for_table_1.dta")
```

```{r generate_table_1}

var_label_list <- list(age = "Age at listing, yrs",
                       female = "Female",
                       race = "Race",
                       History_of_Smoking = "Smoking history",
                       Education_Status = "Education Status",
                       bmi = "BMI, $kg/m^2$",
                       blood_type = "Blood Type",
                       diabetes = "Diabetes",
                       functional_status= "Functional Status",
                      cardiac_index = "Cardiac index,$ml/kg/m^2$",
                      pcwp = "PCWP, mm Hg",
                      payor = "Payor",
                      Working = "Working for income",
                      Diagnosis = "Diagnosis",
                      treatment = "Treatment")

labelled::var_label(for_table_1) <- var_label_list
labelled::var_label(for_table_1)

myvars <- names(for_table_1 %>% dplyr::select(-PX_ID, -policy))

table1 <- CreateTableOne(vars = myvars, strata ="policy", data = for_table_1)

table1_1 <- print(table1, padColnames = TRUE, varLabels = TRUE, dropEqual = TRUE, noSpaces = FALSE, catDigits = 0,  contDigits = 0, test = FALSE, dropEqual== TRUE, printToggle = FALSE)

```

# Adding CI to table 1
```{r categorical variables CI}
#period names
pre_policy_period <-  "Dec 2016 - Feb 2018 (Pre-Policy)"
post_policy_period <- "Dec 2018 - Feb 2020 (Post-policy)"
pre_total <- for_table_1 %>% 
  filter(policy == !! pre_policy_period) %>% 
  nrow()
post_total <- for_table_1 %>% 
  filter(policy == !! post_policy_period) %>% 
  nrow()

#race

race <- for_table_1 %>% 
  select(race, policy) %>% 
  group_by(race, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>% 
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(race, pre, post, change)
  
race <- for_table_1 %>% 
  select(race, policy) %>% 
  group_by(race, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>% 
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>% 
  ungroup() %>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(race, pre, post, change)

#female
female <- for_table_1 %>% 
  select(female, policy) %>% 
  group_by(female, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>% 
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>% 
  ungroup() %>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(female, pre, post, change)

#smoking
History_of_Smoking <- for_table_1 %>% 
  select(History_of_Smoking, policy) %>% 
  group_by(History_of_Smoking, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre))%>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(History_of_Smoking, pre, post, change) 

Working <- for_table_1 %>% 
  select(Working, policy) %>% 
  group_by(Working, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre),
         post = ifelse(is.na(post),0,post))%>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(Working, pre, post, change) 

Education_Status <- for_table_1 %>% 
  select(Education_Status, policy) %>% 
  group_by(Education_Status, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre),
         post = ifelse(is.na(post),0,post))%>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(Education_Status, pre, post, change) 

blood_type <- for_table_1 %>% 
  select(blood_type, policy) %>% 
  group_by(blood_type, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre),
         post = ifelse(is.na(post),0,post))%>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(blood_type, pre, post, change) 

Diagnosis <- for_table_1 %>% 
  select(Diagnosis, policy) %>% 
  group_by(Diagnosis, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre),
         post = ifelse(is.na(post),0,post))%>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(Diagnosis, pre, post, change) 

diabetes <- for_table_1 %>% 
  select(diabetes, policy) %>% 
  group_by(diabetes, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre),
         post = ifelse(is.na(post),0,post))%>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(diabetes, pre, post, change) 

functional_status <- for_table_1 %>% 
  select(functional_status, policy) %>% 
  group_by(functional_status, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre),
         post = ifelse(is.na(post),0,post))%>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(functional_status, pre, post, change) 

payor <- for_table_1 %>% 
  select(payor, policy) %>% 
  group_by(payor, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre),
         post = ifelse(is.na(post),0,post))%>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(payor, pre, post, change) 

library(gridBase)
library(gridExtra)
library(ggplot2)

Table_race <- tableGrob(race)
Table_female <- tableGrob(female)
Table_smoking <- tableGrob(History_of_Smoking)
Table_work <- tableGrob(Working)
Table_edu <- tableGrob(Education_Status)
Table_blood <- tableGrob(blood_type)
Table_diag <- tableGrob(Diagnosis)
Table_diab <- tableGrob(diabetes)
Table_fs <- tableGrob(functional_status)
Table_payor<- tableGrob(payor)

pdf("file.pdf", height = 12, width = 20)
grid.arrange(Table_race, Table_female, Table_smoking, Table_work, Table_edu, Table_blood, Table_diag, Table_diab, Table_fs, Table_payor, ncol = 2, nrow = 5)
dev.off()

```


```{r raw mean difference for continuous variables}
bmi <- for_table_1 %>% 
  select(bmi, policy) %>% 
  mutate(bmi = as.numeric(bmi)) %>% 
  group_by(policy) %>% 
  summarise(mean_bmi = mean(bmi,na.rm=TRUE), sd = sd(bmi, na.rm= TRUE),count = n()) %>%
  mutate(mean_bmi = round(mean_bmi,2),
         sd  = round(sd,2)) %>% 
  ungroup() %>%
  mutate(diff = round(mean_bmi - lag(mean_bmi, default = mean_bmi[1]),2),
         var = (sd^2)/count,
         se_diff = round(sqrt(sum(var)),2),
         policy = ifelse(policy == "Dec 2016 - Feb 2018 (Pre-Policy)", "pre", "post")
         )  %>% 
  filter(policy =="post") %>% 
  select(diff,se_diff) %>% 
  mutate(low_ci = round(diff - 1.96*se_diff,2),
         up_ci = round(diff + 1.96*se_diff,2),
         ci = paste0("[", comma(low_ci), " to ", comma(up_ci), "]"),
         change <- paste0(diff, " ", ci)
         )

age <- for_table_1 %>% 
  select(age, policy) %>% 
  mutate(age = as.numeric(age)) %>% 
  group_by(policy) %>% 
  summarise(mean_age = mean(age,na.rm=TRUE), sd = sd(age, na.rm= TRUE),count = n()) %>%
  ungroup() %>%
  mutate(diff = round(mean_age - lag(mean_age, default = mean_age[1]),1),
         var = (sd^2)/count,
         se_diff = round(sqrt(sum(var)),1),
         policy = ifelse(policy == "Dec 2016 - Feb 2018 (Pre-Policy)", "pre", "post")
         )  %>% 
  filter(policy =="post") %>% 
  select(diff,se_diff) %>% 
  mutate(low_ci = round(diff - 1.96*se_diff,1),
         up_ci = round(diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), " to ", comma(up_ci), "]"),
         change <- paste0(diff, " ", ci)
         )


eGFR <- for_table_1 %>% 
  select(eGFR, policy) %>% 
  mutate(eGFR = as.numeric(eGFR)) %>% 
  group_by(policy) %>% 
  summarise(mean_eGFR = mean(eGFR,na.rm=TRUE), sd = sd(eGFR, na.rm= TRUE),count = n()) %>%
  ungroup() %>%
  mutate(diff = round(mean_eGFR - lag(mean_eGFR, default = mean_eGFR[1]),1),
         var = (sd^2)/count,
         se_diff = round(sqrt(sum(var)),1),
         policy = ifelse(policy == "Dec 2016 - Feb 2018 (Pre-Policy)", "pre", "post")
         )  %>% 
  filter(policy =="post") %>% 
  select(diff,se_diff) %>% 
  mutate(low_ci = round(diff - 1.96*se_diff,1),
         up_ci = round(diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), " to ", comma(up_ci), "]"),
         change <- paste0(diff, " ", ci)
         )


ci <- for_table_1 %>%   
  select(cardiac_index, policy) %>%  
  mutate(cardiac_index = as.numeric(cardiac_index)) %>% 
  group_by(policy) %>%   
  summarise(mean_ci = mean(cardiac_index,na.rm=TRUE), sd = sd(cardiac_index, na.rm= TRUE),count = n()) %>%
  ungroup() %>%
  mutate(mean_ci = round(mean_ci, 2)) %>% 
  mutate(diff = round(mean_ci - lag(mean_ci, default = mean_ci[1]),2),
         var = (sd^2)/count,
         se_diff = round(sqrt(sum(var)),2),
         policy = ifelse(policy == "Dec 2016 - Feb 2018 (Pre-Policy)", "pre", "post")
         )  %>% 
  filter(policy =="post") %>% 
  select(diff,se_diff) %>% 
  mutate(low_ci = round(diff - 1.96*se_diff,2),
         up_ci = round(diff + 1.96*se_diff,2),
         ci = paste0("[", comma(low_ci), " to ", comma(up_ci), "]"),
         change <- paste0(diff, " ", ci))


pcwp <- for_table_1 %>%   
  select(pcwp, policy) %>%  
  mutate(pcwp = as.numeric(pcwp)) %>% 
  group_by(policy) %>%   
  summarise(mean_pcwp = mean(pcwp,na.rm=TRUE), sd = sd(pcwp, na.rm= TRUE),count = n()) %>%
  ungroup() %>%
  mutate(diff = round(mean_pcwp - lag(mean_pcwp, default = mean_pcwp[1]),1),
         var = (sd^2)/count,
         se_diff = round(sqrt(sum(var)),1),
         policy = ifelse(policy == "Dec 2016 - Feb 2018 (Pre-Policy)", "pre", "post")
         )  %>% 
  filter(policy =="post") %>% 
  select(diff,se_diff) %>% 
  mutate(low_ci = round(diff - 1.96*se_diff,1),
         up_ci = round(diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), " to ", comma(up_ci), "]"),
         change <- paste0(diff, " ", ci))


treatment <- for_table_1 %>% 
  select(treatment, policy) %>% 
  group_by(treatment, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>% 
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(treatment, pre, post, change)
  
treatment <- for_table_1 %>% 
  select(treatment, policy) %>% 
  group_by(treatment, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>% 
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>% 
  ungroup() %>% 
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) %>% 
  select(treatment, pre, post, change)
```

```{r Table 1 Updated}
d90_outcome <- final_sample %>% 
  filter(is.na(policy) == FALSE) %>%
  filter(status != "Status 5") %>%  
  dplyr::select(PX_ID,policy,treatment, status, transplanted, list_date, REC_TX_DT,CAN_DEATH_DT) %>% 
  mutate(d_tx = as.Date(REC_TX_DT) - as.Date(list_date),
         d_dx = as.Date(CAN_DEATH_DT) - as.Date(list_date)) %>% 
  mutate(d90 = case_when(
    is.na(d_tx) == FALSE & d_tx <= 90 ~ "transplanted",
    is.na(d_dx) == FALSE & d_dx <= 90 ~ "dead",
    d_tx == 0 ~ "dead",
    d_dx == 0 ~ "transplanted"
  )) %>% 
  mutate(d90 = ifelse(is.na(d90) == TRUE, "censored", d90)) %>% 
  mutate(dftime = case_when(
    d90 == "transplanted" ~ d_tx,
    d90 == "dead" ~ d_dx,
    d90 == "censored" ~ 90
  )) %>% 
  mutate(event_date = case_when(
    is.na(d_tx) == FALSE & d_tx <= 90 ~ as.Date(REC_TX_DT),
    is.na(d_dx) == FALSE & d_dx <= 90 ~ as.Date(CAN_DEATH_DT),
    d_tx == 0 ~ as.Date(REC_TX_DT),
    d_dx == 0 ~ as.Date(CAN_DEATH_DT),
    d90 == "censored" ~ as.Date(as.Date(list_date) + 90)
  )) %>% 
  mutate(half = case_when(
    policy == "Dec 2016 - Feb 2018 (Pre-Policy)" & as.Date(list_date) <= as.Date(as.Date("2016-12-01") + 227) ~ "1",
    policy == "Dec 2016 - Feb 2018 (Pre-Policy)" & as.Date(list_date) > as.Date(as.Date("2016-12-01") + 227) ~ "2",
    policy == "Dec 2018 - Feb 2020 (Post-policy)" & as.Date(list_date) <= as.Date(as.Date("2018-12-01") + 227) ~ "1",
    policy == "Dec 2018 - Feb 2020 (Post-policy)" & as.Date(list_date) > as.Date(as.Date("2018-12-01") + 227) ~ "2",
  ))

which_half <- d90_outcome %>% 
  select(PX_ID,half)

final_sample2 <- final_sample %>% 
  left_join(which_half,by="PX_ID")

#count raw number and proportion of transplantation in 90 days
response <- d90_outcome %>% 
  group_by(policy,d90) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  spread(key = policy, value = count)
  
write.csv(final_sample2, "C:/Users/julia/SRP/SRP/Codes/SAF SAS files/final_sample_half.csv")

pre_total = 4472
post_total = 4498
outcome <- response %>% 
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  mutate(prop_pre = pre/pre_total,
         prop_post = post/post_total,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/pre_total)+
             (prop_post *(1-prop_post)/post_total)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) #%>% 
  select(d90, pre, post, change)
  
#make another data file with candidates parsed in halves based on listing date
  
```
