---
title: Result section for Geographic Variations in Response to New Heart Allocation Policy*
author: "Julia Ran, William Parker"
output:
  rmdformats::material:
    highlight: tango
  html_document:
    theme: cosmo
    toc: yes
---

```{r setup, include=FALSE, echo= FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

require("knitr")
opts_knit$set(root.dir = "C:/Users/julia/Google Drive/Pritzker/SRP/Codes/SAF SAS files")

```

```{r load libraries, echo=FALSE}
library("ggplot2")  
library("tidyverse")
library("haven")
library("rmdformats")
library("ggthemes")
library("dplyr")
```

# Load files
```{r load file, echo=FALSE}
final_sample <- read_csv("C:/Users/julia/Google Drive/Pritzker/SRP/Codes/SAF SAS files/final_sample.csv")

for_table_1 <- read_csv("C:/Users/julia/Google Drive/Pritzker/SRP/Codes/SAF SAS files/for_table_1.csv")

start_date <- as.Date("2016-12-01")
end_date <- as.Date("2020-02-28")
pre_policy_end_date <- as.Date("2018-02-28")
post_policy_start_date <- as.Date("2018-12-01")
policy_switch_date <- as.Date("2018-10-18")
cand_thor <- read_sas("C:/Users/julia/Google Drive/Pritzker/SRP/Codes/SAF SAS files/cand_thor.sas7bdat", NULL) %>%  
  zap_formats() %>% zap_labels()

init_list <- cand_thor %>% 
   mutate(list_date = CAN_LISTING_DT) %>% 
    filter(list_date >= start_date & list_date <= end_date & WL_ORG == "HR")  %>% 
  #filter(list_date <=pre_policy_end_date | list_date >= post_policy_start_date) %>% 
  mutate(status = CAN_INIT_STAT, 
         center = CAN_LISTING_CTR_ID,
         date_start = list_date) %>% 
  filter(CAN_AGE_AT_LISTING >17) 


modelhigh_output<- read.csv("C:/Users/julia/Google Drive/Pritzker/SRP/Codes/data_codes/high_stata_output.csv",sep="\t", header=TRUE)

model_highstatus <- modelhigh_output %>% 
  select(ctr_cd, policy_factor, high_status)

modelhigh_output <- modelhigh_output %>% 
  select(-high_status)

CMS_adjusted <- read_csv("C:/Users/julia/Google Drive/Pritzker/SRP/Codes/data_codes/CMS_adjusted_rates.csv")

CMS_full <- read_csv("C:/Users/julia/Google Drive/Pritzker/SRP/Codes/data_codes/figure_2_CMS.csv")

logit2prob <- function(logit){
  odds <- exp(-logit)
  prob <- 1 / (1 + odds)
  return(prob)
}

model_high_clean <- modelhigh_output %>% 
  distinct(ctr_cd, policy_factor, .keep_all = TRUE) %>% 
  group_by(ctr_cd) %>% 
  spread(key = policy_factor, value = N_ctr) %>% 
  ungroup() %>% 
  rename(N_pre = Pre,
         N_post = Post) %>% 
  mutate(
    pct_pre = prob_pre*100,
    pct_post = prob_post*100,
    diff = (prob_post - prob_pre)*100,
    se_pre = sqrt(prob_pre*(1-prob_pre)/N_pre),
    se_post = sqrt(prob_post*(1-prob_post)/N_post),
    lower_ci_pre = pct_pre -1.96*se_pre*100,
    upper_ci_pre = pct_pre +1.96*se_pre*100,
    lower_ci_post = pct_post - 1.96*se_post*100,
    upper_ci_post = pct_post +1.96*se_post*100,
    lower_ci_pre = ifelse(lower_ci_pre<0,0,lower_ci_pre),
    lower_ci_post = ifelse(lower_ci_post<0,0,lower_ci_post)
  ) %>% 
  mutate(
    diff_sig = prob_post - prob_pre,
    se_diff = sqrt(prob_pre*(1-prob_pre)/N_pre+prob_post*(1-prob_post)/N_post),
    t_diff = diff_sig/se_diff,
    sig = ifelse(t_diff >1.96, as.character(upper_ci_post + 5), ""),
    sig1 = as.numeric(sig)
  )
```

# Functions for formating numbers
```{r formatting functions, echo=FALSE}
comma <- function(x){
  case_when(
    abs(x) > 10 ~ format(x, digits = 0, big.mark = ","),
    abs(x) > 1 ~ format(x, digits = 3, big.mark = ","),
    TRUE ~ format(x, digits = 3, big.mark = ",")
  )
    
}

comma_p <- function(x){
  case_when(
    x < 0.001 ~ "p<0.001",
    x < 0.01 ~ paste0("p=" ,format(x, digits = 1, big.mark = ",")),
    TRUE ~ paste0("p=" ,format(x, digits = 3, big.mark = ","))
  )
}
```

# Paragraph 1
```{r sample size, echo=FALSE}
n_total = nrow(init_list)

n_ctr = final_sample %>% 
  filter(!is.na(policy)) %>% 
  filter(status !="Status 5") %>% 
  select(CTR_CD) %>% 
  distinct() %>% 
  nrow()

n_opo = final_sample %>% 
  filter(!is.na(policy)) %>% 
  filter(status !="Status 5") %>% 
  select(SERVED_OPO_CD) %>% 
  distinct() %>% 
  nrow()

n_included = nrow(for_table_1)

n_excluded = n_total - n_included

pre_policy <- for_table_1 %>% 
    filter(policy == "Dec 2016 - Feb 2018 (Pre-Policy)") 

post_policy <- for_table_1 %>% 
    filter(policy == "Dec 2018 - Feb 2020 (Post-policy)")  

n_total_pre = pre_policy %>% 
  nrow()

n_total_post = post_policy %>% 
  nrow()

res_n = prop.test(x = n_total_pre, n = n_included, p = 0.5, correct = FALSE)
```

```{r white, echo=FALSE}
white_pre = pre_policy %>% 
  filter(race == "White") %>% 
  nrow()

pct_white_pre = paste0(comma(100*(white_pre/n_total_pre)),"%")


white_post = post_policy %>% 
  filter(race == "White") %>% 
  nrow()

pct_white_post = paste0(comma(100*(white_post/n_total_post)),"%")

se_white_diff = sqrt(
           ((white_pre/n_total_pre) * (1-(white_pre/n_total_pre))/n_total_pre)+
             ((white_post/n_total_post) *(1-(white_post/n_total_post))/n_total_post))*100

abs_white_diff = round((white_post/n_total_post)*100 - (white_pre/n_total_pre)*100, 1)
low_white_ci = comma(abs_white_diff - 1.96*se_white_diff)
up_white_ci = comma(abs_white_diff + 1.96*se_white_diff)

white_stat = paste0("absolute difference, ", abs_white_diff, "%", "; 95% CI: ", low_white_ci, "% to ", up_white_ci,"%")


```

```{r work, echo=FALSE}
work_pre = pre_policy %>% 
  filter(Working == "Working") %>% 
  nrow()

pct_work_pre = paste0(comma(100*(work_pre/n_total_pre)),"%")


work_post = post_policy %>% 
  filter(Working == "Working") %>% 
  nrow()

pct_work_post = paste0(comma(100*(work_post/n_total_post)),"%")

se_work_diff = sqrt(
           ((work_pre/n_total_pre) * (1-(work_pre/n_total_pre))/n_total_pre)+
             ((work_post/n_total_post) *(1-(work_post/n_total_post))/n_total_post))*100

abs_work_diff = round((work_post/n_total_post)*100 - (work_pre/n_total_pre)*100, 1)
low_work_ci = comma(abs_work_diff - 1.96*se_work_diff)
up_work_ci = comma(abs_work_diff + 1.96*se_work_diff)

work_stat = paste0("absolute difference, ", abs_work_diff, "%", "; 95% CI: ", low_work_ci, "% to ", up_work_ci,"%")

```


```{r smoke, echo=FALSE}
smoke_pre = pre_policy %>% 
  filter(History_of_Smoking == "Smoking history") %>% 
  nrow()

pct_smoke_pre = paste0(round(100*(smoke_pre/n_total_pre),1),"%")


smoke_post = post_policy %>% 
  filter(History_of_Smoking == "Smoking history") %>% 
  nrow()

pct_smoke_post = paste0(round(100*(smoke_post/n_total_post),1),"%")

se_smoke_diff = sqrt(
           ((smoke_pre/n_total_pre) * (1-(smoke_pre/n_total_pre))/n_total_pre)+
             ((smoke_post/n_total_post) *(1-(smoke_post/n_total_post))/n_total_post))*100

abs_smoke_diff = round((smoke_post/n_total_post)*100 - (smoke_pre/n_total_pre)*100, 1)
low_smoke_ci = round(abs_smoke_diff - 1.96*se_smoke_diff,1)
up_smoke_ci = round(abs_smoke_diff + 1.96*se_smoke_diff,1)

smoke_stat = paste0("absolute difference, ", abs_smoke_diff, "%", "; 95% CI: ", low_smoke_ci, "% to ", up_smoke_ci,"%")

```

```{r bmi, echo=FALSE}
bmi <- for_table_1 %>% 
  select(bmi, policy) %>% 
  mutate(bmi = as.numeric(bmi)) %>% 
  group_by(policy) %>% 
  summarise(mean_bmi = mean(bmi,na.rm=TRUE), sd = sd(bmi, na.rm= TRUE),count = n()) %>%
  mutate(mean_bmi = round(mean_bmi,1),
         sd  = round(sd,1)) %>% 
  ungroup() %>%
  mutate(diff = round(mean_bmi - lag(mean_bmi, default = mean_bmi[1]),1),
         var = (sd^2)/count,
         se_diff = round(sqrt(sum(var)),1),
         policy = ifelse(policy == "Dec 2016 - Feb 2018 (Pre-Policy)", "pre", "post")
         )  %>% 
  filter(policy =="post") %>% 
  select(diff,se_diff) %>% 
  mutate(low_ci = round(diff - 1.96*se_diff,1),
         up_ci = round(diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), " to ", comma(up_ci), "]"),
         change <- paste0(diff, " ", ci)
         )

abs_bmi_diff = paste0(bmi$diff)
low_bmi_ci = round(bmi$low_ci,1)
up_bmi_ci = round(bmi$up_ci,1)

bmi_stat = paste0("absolute difference, ", abs_bmi_diff,"; 95% CI: ", low_bmi_ci, " to ", up_bmi_ci)

```

```{r ci, echo=FALSE}
ci <- for_table_1 %>%   
  select(cardiac_index, policy) %>%  
  mutate(cardiac_index = as.numeric(cardiac_index)) %>% 
  group_by(policy) %>%   
  summarise(mean_ci = mean(cardiac_index,na.rm=TRUE), sd = sd(cardiac_index, na.rm= TRUE),count = n()) %>%
  ungroup() %>%
  mutate(mean_ci = mean_ci) %>% 
  mutate(diff = mean_ci - lag(mean_ci, default = mean_ci[1]),
         var = (sd^2)/count,
         se_diff = sqrt(sum(var)),
         policy = ifelse(policy == "Dec 2016 - Feb 2018 (Pre-Policy)", "pre", "post")
         )  %>% 
  filter(policy =="post") %>% 
  select(diff,se_diff) %>% 
  mutate(low_ci = diff - 1.96*se_diff,
         up_ci = diff + 1.96*se_diff)

abs_ci_diff = round(ci$diff,2)
low_ci_ci = round(ci$low_ci,2)
up_ci_ci = round(ci$up_ci,2)

ci_stat = paste0("absolute difference, ", abs_ci_diff,"; 95% CI: ", low_ci_ci, " to ", up_ci_ci)
```

```{r fs, echo=FALSE}
functional_status <- for_table_1 %>% 
  select(functional_status, policy) %>% 
  group_by(functional_status, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre),
         post = ifelse(is.na(post),0,post))%>% 
  mutate(prop_pre = pre/n_total_pre,
         prop_post = post/n_total_post,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/n_total_pre)+
             (prop_post *(1-prop_post)/n_total_post)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`) 


abs_fs_diff = paste0(functional_status$abs_pct_diff[3])
low_fs_ci = paste0(functional_status$low_ci[3])
up_fs_ci = paste0(functional_status$up_ci[3])

fs_stat = paste0("absolute difference in % with severe impairment, ", abs_fs_diff, "%", "; 95% CI: ", low_fs_ci, "% to ", up_fs_ci,"%")

```

```{r fs_missing, echo=FALSE}
abs_fsm_diff = paste0(round(functional_status$abs_pct_diff[4],1))
low_fsm_ci = paste0(round(functional_status$low_ci[4],1))
up_fsm_ci = paste0(round(functional_status$up_ci[4],1))

fsm_stat = paste0("absolute difference, ", abs_fsm_diff, "%", "; 95% CI: ", low_fsm_ci, "% to ", up_fsm_ci,"%")

```

```{r db_missing, echo=FALSE}
diabetes <- for_table_1 %>% 
  select(diabetes, policy) %>% 
  group_by(diabetes, policy) %>% 
  summarise(count = n()) %>% 
  spread(key = policy, value = count) %>%
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>%
  ungroup() %>% 
  mutate(pre = ifelse(is.na(pre), 0, pre),
         post = ifelse(is.na(post),0,post))%>% 
  mutate(prop_pre = pre/n_total_pre,
         prop_post = post/n_total_post,
         se_diff = sqrt(
           (prop_pre * (1-prop_pre)/n_total_pre)+
             (prop_post *(1-prop_post)/n_total_post)
           )*100,
         pct_pre = prop_pre*100,
         pct_post = prop_post*100,
         abs_pct_diff = round(pct_post - pct_pre, 1),
         low_ci = round(abs_pct_diff - 1.96*se_diff,1),
         up_ci = round(abs_pct_diff + 1.96*se_diff,1),
         ci = paste0("[", comma(low_ci), "% to ", comma(up_ci), "%]"),
         change <- paste0(abs_pct_diff, " ", ci)
         ) %>% 
  rename(change = `change <- paste0(abs_pct_diff, " ", ci)`)

abs_dbm_diff = paste0(diabetes$abs_pct_diff[3])
low_dbm_ci = paste0(diabetes$low_ci[3])
up_dbm_ci = paste0(diabetes$up_ci[3])

dbm_stat = paste0("absolute difference, ", abs_dbm_diff, "%", "; 95% CI: ", low_dbm_ci, "% to ", up_dbm_ci,"%")

```

```{r pcwp}
pcwp <- for_table_1 %>%   
  select(pcwp, policy) %>%  
  mutate(pcwp = as.numeric(pcwp)) %>% 
  group_by(policy) %>%   
  summarise(mean_pcwp = mean(pcwp,na.rm=TRUE), sd = sd(pcwp, na.rm= TRUE),count = n()) %>%
  ungroup() %>%
  mutate(mean_pcwp = mean_pcwp) %>% 
  mutate(diff = mean_pcwp - lag(mean_pcwp, default = mean_pcwp[1]),
         var = (sd^2)/count,
         se_diff = sqrt(sum(var)),
         policy = ifelse(policy == "Dec 2016 - Feb 2018 (Pre-Policy)", "pre", "post")
         )  %>% 
  filter(policy =="post") %>% 
  select(diff,se_diff) %>% 
  mutate(low_ci = diff - 1.96*se_diff,
         up_ci = diff + 1.96*se_diff)

abs_pcwp_diff = round(pcwp$diff,2)
low_pcwp_ci = round(pcwp$low_ci,2)
up_pcwp_ci = round(pcwp$up_ci,2)

pcwp_stat = paste0("absolute difference, ", abs_pcwp_diff,"; 95% CI: ", low_pcwp_ci, " to ", up_pcwp_ci)
```

A total of `r comma(n_total)` active, adult heart-only candidates were listed during the study period. `r n_excluded` candidates were excluded from analysis using exclusion criteria listed in Figure 1. The total size of the study population was `r n_included` from `r n_ctr` transplant centers within `r n_opo` Organ Procurement Organizations (OPO). The number of patients from the pre-policy period (Dec 1, 2016 to Feb 28, 2018) and post-policy period (Dec 1, 2018 to Feb 28, 2020) were comparable (`r n_total_pre` vs `r n_total_post`, `r comma_p(res_n$p.value)` by one-proportion Z-test). Table 1 compares candidate characteristics between the pre-policy and post-policy cohort. Overall, the socioeconomic and clinical characteristics of listed candidates remained largely unchanged across policy periods. Candidates listed in the post-policy period  exhibited slightly worse cardiac index (`r ci_stat`), and worse functional status (`r fs_stat`). Mean pulmonary capillary wedge pressure was comparable between the two policy periods (`r pcwp_stat`)


# Paragraph 2
```{r status, echo= FALSE}
status <- final_sample %>% 
  filter(!is.na(policy),
         status !="Status 5") %>% 
  select(policy, status) %>% 
  group_by(status, policy) %>% 
  count() %>% 
  spread(key = policy, value = n) %>% 
  rename(pre = `Dec 2016 - Feb 2018 (Pre-Policy)`,
         post = `Dec 2018 - Feb 2020 (Post-policy)`) %>% 
  mutate(pct_s_pre = 100*pre/n_total_pre,
         pct_s_post = 100*post/n_total_post,
         diff = pct_s_post - pct_s_pre,
         se_s_diff = sqrt(
           (pre/n_total_pre) * (1-(pre/n_total_pre))/n_total_pre +
             ((post/n_total_post) *(1-(post/n_total_post))/n_total_post))*100)

abs_s1 = status$diff[1]
abs_s2 = status$diff[2]
abs_s3 = status$diff[3]
abs_s4 = status$diff[4]
abs_s6 = status$diff[5]

low_s1_ci = abs_s1 - 1.96*status$se_s_diff[1]
low_s2_ci = abs_s2 - 1.96*status$se_s_diff[2]
low_s3_ci = abs_s3 - 1.96*status$se_s_diff[3]
low_s4_ci = abs_s4 - 1.96*status$se_s_diff[4]
low_s6_ci = abs_s6 - 1.96*status$se_s_diff[5]

up_s1_ci = abs_s1 + 1.96*status$se_s_diff[1]
up_s2_ci = abs_s2 + 1.96*status$se_s_diff[2]
up_s3_ci = abs_s3 + 1.96*status$se_s_diff[3]
up_s4_ci = abs_s4 + 1.96*status$se_s_diff[4]
up_s6_ci = abs_s6 + 1.96*status$se_s_diff[5]

n_s2_pre = paste0(comma(status$pre[2]))
p_s2_pre = paste0(round(status$pct_s_pre[2],1), "%")
n_s2_post =paste0(comma(status$post[2]))
p_s2_post = paste0(round(status$pct_s_post[2],1), "%")
s2_stat = paste0("absolute difference, ", round(abs_s2,1), "%", "; 95% CI: ", round(low_s2_ci,1), "% to ", round(up_s2_ci,1),"%")



n_s1_pre = paste0(comma(status$pre[1]))
p_s1_pre = paste0(round(status$pct_s_pre[1],1), "%")
n_s1_post =paste0(comma(status$post[1]))
p_s1_post = paste0(round(status$pct_s_post[1],1), "%")
s1_stat = paste0("absolute difference, ", round(abs_s1,1), "%", "; 95% CI: ", round(low_s1_ci,1), "% to ", round(up_s1_ci,1),"%")


n_s3_pre = paste0(comma(status$pre[3]))
p_s3_pre = paste0(round(status$pct_s_pre[3],1), "%")
n_s3_post =paste0(comma(status$post[3]))
p_s3_post = paste0(round(status$pct_s_post[3],1), "%")
s3_stat = paste0("absolute difference, ", comma(abs_s3), "%", "; 95% CI: ", comma(low_s3_ci), "% to ", comma(up_s3_ci),"%")


n_s4_pre = paste0(comma(status$pre[4]))
p_s4_pre = paste0(round(status$pct_s_pre[4],1), "%")
n_s4_post =paste0(comma(status$post[4]))
p_s4_post = paste0(round(status$pct_s_post[4],1), "%")
s4_stat = paste0("absolute difference, ", comma(abs_s4), "%", "; 95% CI: ", comma(low_s4_ci), "% to ", comma(up_s4_ci),"%")


n_s6_pre = paste0(comma(status$pre[5]))
p_s6_pre = paste0(round(status$pct_s_pre[5],1), "%")
n_s6_post =paste0(comma(status$post[5]))
p_s6_post = paste0(round(status$pct_s_post[5],1), "%")
s6_stat = paste0("absolute difference, ", round(abs_s6,1), "%", "; 95% CI: ", round(low_s6_ci,1), "% to ", round(up_s6_ci,1),"%")

```

After applying the new status justification to candidates listed in pre-policy period, the expected status distribution was `r p_s1_pre` Status 1, `r p_s2_pre` Status 2, `r p_s3_pre` Status 3, `r p_s4_pre` Status 4, and `r p_s6_pre` Status 6. The observed status distribution after implementing the new policy was `r p_s1_post` Status 1, `r p_s2_post` Status 2, `r p_s3_post` Status 3, `r p_s4_post` Status 4, and `r p_s6_post` Status 6. The post-policy distribution was significantly different from the pre-policy distribution (p < 0.001 by chi-squared test). There were more high-priority listings than expected, with +`r round(abs_s1,1)` Status 1 (95% CI: `r round(low_s1_ci,1)` to `r round(up_s1_ci,1)`) and `r round(abs_s2,1)` Status 2 (95% CI: `r round(low_s2_ci,1)` to `r round(up_s2_ci,1)`). In contrast, there were fewer Status 3 (`r s3_stat`), Status 4 (`r s4_stat`), and Status 6 listings (`r s6_stat`). Figure 1 displays the monthly changes in the number of candidates assigned to each priority status across the two policy periods.


# Justification for high-status listing in the post-policy period

```{r exception, echo=FALSE}
exception <- final_sample %>% 
  filter(!is.na(policy),
         status !="Status 5") %>% 
  select(policy, status, treatment,status_1a) %>% 
  mutate(status_total = case_when(
    status_1a == "TRUE" & policy == "Dec 2016 - Feb 2018 (Pre-Policy)" ~ "Status 1A",
    status_1a == "FALSE" & policy == "Dec 2016 - Feb 2018 (Pre-Policy)" ~ "Other", 
    status_1a == "FALSE" & policy != "Dec 2016 - Feb 2018 (Pre-Policy)" ~ as.character(status)
  )) %>% 
  select(policy, status_total, treatment) #%>% 
  group_by(status_total, treatment) %>% 
  summarise(count = n()) %>% 
  mutate(pct = round(count/sum(count)*100,0)) %>% 
  ungroup()

#pre-policy Status 1A treatment breakdown: ECMO, IABP, LVAD, other MCS, exception
n_exp_1a <- exception$count[exception$status_total =="Status 1A" & exception$treatment == "Exception"]
p_exp_1a <- exception$pct[exception$status_total =="Status 1A" & exception$treatment == "Exception"]

n_mcs_1a <- exception$count[exception$status_total =="Status 1A" & exception$treatment == "Other MCS"]
p_mcs_1a <- exception$pct[exception$status_total =="Status 1A" & exception$treatment == "Other MCS"]

n_ecmo_1a<- exception$count[exception$status_total =="Status 1A" & exception$treatment == "ECMO"]
p_ecmo_1a <- exception$pct[exception$status_total =="Status 1A" & exception$treatment == "ECMO"]

n_lvad_1a<- exception$count[exception$status_total =="Status 1A" & exception$treatment == "LVAD"]
p_lvad_1a <- exception$pct[exception$status_total =="Status 1A" & exception$treatment == "LVAD"]

n_iabp_1a<- exception$count[exception$status_total =="Status 1A" & exception$treatment == "IABP"]
p_iabp_1a <- exception$pct[exception$status_total =="Status 1A" & exception$treatment == "IABP"]

#Status 1 treatment breakdown: ECMO, exception, other MCS
n_ecmo_1<- exception$count[exception$status_total =="Status 1" & exception$treatment == "ECMO"]
p_ecmo_1 <- exception$pct[exception$status_total =="Status 1" & exception$treatment == "ECMO"]

n_exp_1 <- exception$count[exception$status_total =="Status 1" & exception$treatment == "Exception"]
p_exp_1 <- exception$pct[exception$status_total =="Status 1" & exception$treatment == "Exception"]

n_mcs_1 <- exception$count[exception$status_total =="Status 1" & exception$treatment == "Other MCS"]
p_mcs_1 <- exception$pct[exception$status_total =="Status 1" & exception$treatment == "Other MCS"]

#Status 2 treatment breakdown: IABP, LVAD, other MCS, exception

n_iabp_2<- exception$count[exception$status_total =="Status 2" & exception$treatment == "IABP"]
p_iabp_2 <- exception$pct[exception$status_total =="Status 2" & exception$treatment == "IABP"]

n_lvad_2<- exception$count[exception$status_total =="Status 2" & exception$treatment == "LVAD"]
p_lvad_2 <- exception$pct[exception$status_total =="Status 2" & exception$treatment == "LVAD"]

n_mcs_2 <- exception$count[exception$status_total =="Status 2" & exception$treatment == "Other MCS"]
p_mcs_2 <- exception$pct[exception$status_total =="Status 2" & exception$treatment == "Other MCS"]

n_exp_2 <- exception$count[exception$status_total =="Status 2" & exception$treatment == "Exception"]
p_exp_2 <- exception$pct[exception$status_total =="Status 2" & exception$treatment == "Exception"]

exp_diff <- round((n_exp_1 + n_exp_2 - n_exp_1a)/n_exp_1a, 1)

```

In the pre-policy cohort, `r n_ecmo_1a` (`r p_ecmo_1a`%) candidates listed in Status 1A received ECMO, `r n_iabp_1a` (`r p_iabp_1a`%) received IABP, `r n_lvad_1a` (`r p_lvad_1a` %) received LVAD, and `r n_mcs_1a` (`r p_mcs_1a`%) received other MCS. `r n_exp_1a` (`r p_exp_1a`%) of Status 1A listings were justified by exception requests. 

In the post-policy cohort, the observed Status 1 listings were justified by `r n_ecmo_1` (`r p_ecmo_1`%) ECMO, `r n_mcs_1` (`r p_mcs_1`%) MCS and `r n_exp_1` (`r p_exp_1`%) exception requests. Status 2 listings were justified by `r n_lvad_2` (`r p_lvad_2`%) LVAD, `r n_iabp_2`IABP, `r n_mcs_2` (`r p_mcs_2`%) other MCS, and `r n_exp_2` (`r p_exp_2`%) exception requests. The observed number of exception requests among Status 1 and 2 listings exceeded that of Status 1A by `r exp_diff` times. 

# Estimated Probability of High Priority Status Listing Across Policy Periods (without CMS approach)

```{r diff_change, echo = FALSE}

range_diff = model_high_clean %>% 
  select(ctr_cd, diff) %>% 
  arrange(diff) %>% 
  mutate(index = row_number()) 

min_diff = round(range_diff$diff[1],1)
max_diff = round(range_diff$diff[length(range_diff$diff)],1)

iqr_diff = paste0("Inter-quartile range: ", round(quantile(range_diff$diff,0.25)),  "% to %", round(quantile(range_diff$diff,0.75),1)) 

n_ctr_sig <- model_high_clean %>% 
  select(ctr_cd, sig) %>% 
  mutate(sig_count = ifelse(sig!="", as.numeric("1"), as.numeric("0"))) %>% 
  summarise(count = sum(sig_count))


n_ctr_sig = paste0(n_ctr_sig)

p_ctr_sig = round(as.numeric(n_ctr_sig)/n_ctr*100,1)


```

```{r pre high_status range, echo = FALSE}
range_pre = model_high_clean %>% 
  select(ctr_cd, pct_pre) %>% 
  arrange(pct_pre)

min_pre = round(range_pre$pct_pre[1],1)

max_pre = round(range_pre$pct_pre[length(range_pre$pct_pre)],1)

iqr_pre = paste0("IQR: ", round(quantile(range_pre$pct_pre,0.25),1),  "% to %", round(quantile(range_pre$pct_pre,0.75),1)) 

pre_range = paste0("min = ", min_pre, "%, max = ", max_pre,  "%, ", iqr_pre)
```

```{r post high_status range, echo = FALSE}
range_post = model_high_clean %>% 
  select(ctr_cd, pct_post) %>% 
  arrange(pct_post)

min_post = round(range_post$pct_post[1],1)

max_post = round(range_post$pct_post[length(range_post$pct_post)],1)

iqr_post = paste0("IQR: ", round(quantile(range_post$pct_post,0.25),1),  "% to ", round(quantile(range_post$pct_post,0.75),1), "%") 

post_range = paste0("min = ", min_post, "%, max = ", max_post,  "%, ", iqr_post)
```

```{r Number of centers >10% and >20% post policy}
N_10 <- model_high_clean %>% 
  select(ctr_cd, diff) %>% 
  filter(diff>10) %>% 
  nrow()

N_10 <- as.numeric(paste0(N_10))

p_10 <- paste0(round(N_10/n_ctr,2)*100,"%")



N_20 <- model_high_clean %>% 
  select(ctr_cd, diff) %>% 
  filter(diff>20) %>% 
  nrow()

N_20 <- as.numeric(paste0(N_20))

p_20 <- paste0(round(N_20/n_ctr,2)*100,"%")

```


# Estimated Probability of High Priority Status Listing Across Policy Periods (WITH CMS approach)

```{r diff_change, echo = FALSE}
range_diff = CMS_adjusted %>% 
  select(ctr_cd, policy_factor, adjusted_rate) %>% 
  group_by(ctr_cd) %>% 
  spread(key = policy_factor, value = adjusted_rate) %>% 
  ungroup()%>%
  mutate(diff = (Post - Pre)*100) %>% 
  arrange(diff) %>% 
  mutate(index = row_number()) 

min_diff = round(range_diff$diff[1],1)
max_diff = round(range_diff$diff[length(range_diff$diff)],1)

iqr_diff = paste0("Inter-quartile range: ", round(quantile(range_diff$diff,0.25)),  "% to %", round(quantile(range_diff$diff,0.75),1)) 

n_ctr_sig <- CMS_full %>% 
  select(ctr_cd, policy_factor, sig1) %>% 
  filter(policy_factor ==2) %>% 
  filter(!is.na(sig1)) %>% 
  summarise(count = n())


n_ctr_sig = paste0(n_ctr_sig)

p_ctr_sig = paste0(round(as.numeric(n_ctr_sig)/n_ctr*100,1))


```

```{r pre high_status range, echo = FALSE}
range_pre = CMS_full %>% 
  select(ctr_cd, policy_factor, adjusted_rate) %>% 
  filter(policy_factor == 1) %>% 
  mutate(adjusted_rate = adjusted_rate*100) %>% 
  arrange(adjusted_rate)

min_pre = round(range_pre$adjusted_rate[1],1)

max_pre = round(range_pre$adjusted_rate[length(range_pre$adjusted_rate)],1)

iqr_pre = paste0("IQR: ", round(quantile(range_pre$adjusted_rate,0.25),1),  "% to %", round(quantile(range_pre$adjusted_rate,0.75),1)) 

pre_range = paste0("min = ", min_pre, "%, max = ", max_pre,  "%, ", iqr_pre)
```

```{r post high_status range, echo = FALSE}
range_post = CMS_full %>% 
  select(ctr_cd, policy_factor, adjusted_rate) %>% 
  filter(policy_factor == 2) %>% 
  mutate(adjusted_rate = adjusted_rate *100) %>% 
  arrange(adjusted_rate)

min_post = round(range_post$adjusted_rate[1],1)

max_post = round(range_post$adjusted_rate[length(range_post$adjusted_rate)],1)

iqr_post = paste0("IQR: ", round(quantile(range_post$adjusted_rate,0.25),1),  "% to ", round(quantile(range_post$adjusted_rate,0.75),1), "%") 

post_range = paste0("min = ", min_post, "%, max = ", max_post,  "%, ", iqr_post)
```

```{r Number of centers >10% and >20% post policy}
N_10 <- range_diff %>% 
  select(ctr_cd, diff) %>% 
  filter(diff>0.1) %>% 
  nrow()

N_10 <- as.numeric(paste0(N_10))

p_10 <- paste0(round(N_10/n_ctr,2)*100,"%")



N_20 <- range_diff %>% 
  select(ctr_cd, diff) %>% 
  filter(diff>0.2) %>% 
  nrow()

N_20 <- as.numeric(paste0(N_20))

p_20 <- paste0(round(N_20/n_ctr,2)*100,"%")

```

On average, in the pre-policy cohort 7% of candidates were met high-priority status criteria (Status 1 & 2) compared to the 25% candidates listed at high-priority in the post-policy period (p <0.001). In the multilevel logistic regression accounting for center effects, the odds of listing at high-priority status was 423% higher in the post-policy period than expected (OR 5.23, 95% CI 4.26-6.42, Table 2). After controlling for candidate level variables, the odds were 534% higher (adjusted OR 6.34, 5.08-7.91%).

After adjusting for changes in candidate-level characteristics, the expected pre-policy  high priority status listing rate varied from `r min_pre`% to `r max_pre`% (`r iqr_pre`). The observed high-status listing rates in the post-policy period varied from `r min_post`%, to `r max_post`% (`r iqr_post`) (Figure 2). A total of `r n_ctr_sig` (`r p_ctr_sig`%) of centers listed significantly more patients at high-priority status than expected (Figure 2). The observed rate of high-priority status listing exceeded expected by dramatically different rates between transplant centers, ranging from +2.2 % to +41.7 %, (Inter-quartile range: 8% to 18.4%).  A total of `r N_10` centers (`r p_10`) listed 10% more candidates at high status than expected and `r N_20` centers (`r p_20`) listed 20% more candidates than expected.


# Association of center and OPO level variables and the change in high-status listing

Figure 3 compares the geographic variation in the rate of high-priority status listing before and after implementing the new heart allocation policy. After adjusting for candidate-level characteristics and weighting by listing volume at constituent transplant centers, the rate of high-priority status listing increased in all OPOs. While the magnitude of the increase differs substantially across OPOs, its geographical pattern is inconspicuous. Areas of the greatest rates of high-priority status listing in the post-policy period not only include large urban areas, such as New York, Chicago, San Francisco, and Los Angeles, but also many OPOs in Midwestern and Southern United States.

Among the three center-level predictors tested, the proportion of Status 1A patients receiving a transplant in the pre-policy period was significantly associated with greater chance of high-priority status listing (Table 2, see Supplement for full models). Transplant centers in OPOs with the highest quartile of pre-policy Status 1A transplant rate (>82%) were 4.18 times more likely to list a patient at high-priority status (aOR: 9.70; 95% CI: 6.65 to 14.13) in the post-policy period than those in OPOs with the lowest quartile (less than 72%) of Status 1A transplant rate (aOR: 5.52; 95% CI: 3.89,7.84). The variance of the random slope (0.28) quantifies the different correlations between policy implementation between transplant centers, and remains statistically significant across the three models.


# Abstract

```{r}
#XX% IABP, YY% ECMO, ZZ% exception, WW% other . 
p_iabp <- (n_iabp_2)/(as.numeric(n_s1_post) + as.numeric(n_s2_post))

print_iabp <- paste0(round(p_iabp*100,0),"%")

p_ecmo <- (n_ecmo_1)/(as.numeric(n_s1_post) + as.numeric(n_s2_post))
print_ecmo <- paste0(round(p_ecmo*100,0),"%")

p_exp <- (n_exp_1 + n_exp_2)/(as.numeric(n_s1_post) + as.numeric(n_s2_post))
print_exp <- paste0(round(p_exp*100,0),"%")

p_other <- 1- (p_iabp + p_ecmo + p_exp)
print_other <- paste0(round(p_other*100,0),"%")

```

The therapies used to justify high-priority status listings were `r print_iabp` IABP, `r print_ecmo`% ECMO, `r print_exp` exception, `r print_other` other . 


```{r chi-square test}
status1 <-  status %>% 
  select(status, post, pct_s_pre) %>% 
  mutate(pct_s_pre = pct_s_pre/100)

chisq.test(status1$post, p = status1$pct_s_pre)


```